---
title: "GLM HW1"
author: "Charlotte Li"
date: "8/18/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part I Binary Outcome 
```{r}
setwd("~/Documents/nyu/1stGradSpring/Generalized Linear Model/dataset")
dat<-read.csv("RELIGION.csv")
dat$relschol <- as.numeric(dat$relschol=="yes")
dat$white <- as.numeric(dat$race=="white")
```
## 1
```{R}
## probability of attending religious school 
sum(dat$relschol==1)/sum(nrow(dat))
## compute the corresponding odds 
sum(dat$relschol==1)/sum(dat$relschol==0)
```
## 2
```{R}
table(dat$white,dat$relschol)
## probability of non-white attend religious school
26/102
## probability of white attend religious school
54/524
```

## 3
```{r}
# odds that non-white students attend religious school
26/76
# odds that white students attend religious school
54/470
```
## 4
the odds ratio that compares the odds of white versus non-white attend religious school
```{r}
(54/470)/(26/76)
```
## 5 
Build a logistic regression to predict "relschol"using variables "white", "attend" and "age", treating the latter two as continuous variables. Report the odds ratio for variable "white"
```{r}
summary(glm(formula=relschol~as.factor(white)+attend+age,data=dat,family="binomial"))
exp(-0.94)
```

## 6 
In one short sentence, explain the meaning of the odds ratio for "white" you reported in question 5: 
--The odds ratio for white students of attending religious school is 0.39 times the odds of non white students attending religious school, holding other variables constant.

## 7 
Report the adjusted odds ratio comparing Non-white students versus White students based one the results from the model in question 5.
```{r}
1/exp(-0.94)
```

## 8 
Further extend the model in previous question by including a quadratic term of age, "agesq". Run a logistic regression now with predictors: white, attend, age and agesq. Choose the answer that is best informed by the results of this model (as compared with the previous one)
```{r}
dat$agesq<-dat$age*dat$age
summary(model8<-glm(formula=relschol~as.factor(white)+attend+age+agesq,data=dat,family="binomial"))
```
 The age of respondent has a quadratic (curvilinear) relationship with the log odds of attending religious school.
 
## 9
```{R}
res <- glm(relschol~as.factor(white+attend) + age +agesq, data=dat,family=binomial(link="logit"))
newdat <- matrix(0, 2,4)
newdat[1,]<-c(0, 5, 45, 45^2)
newdat[2,]<-c(1,5,45,45^2)
newdat<-as.data.frame(newdat)
names(newdat)<-c("white","attend","age", "agesq")
pp <- predict(model8, newdata=newdat, type="response", se.fit=TRUE)
pp
```
For those who attend religious services five days per month (attend=5) and age at 45, what is the predicted probability of having attended a religious school for non-white students:  0.384  with a standard error of  0.069 ; and the predicted probability for white students:  0.192 with a standard error of  0.030

## 10
ABCDEF

## 11 
Run another model, include white, attend and age and agesq, treating "attend" as categorical variable. Based on AIC and BIC, which model fits the data better? Note a smaller AIC indicates better model fit considering the number of predictors used (model degrees of freedom) [hint: in R use AIC(res) to pull out the AIC value of a model. In Stata, use glm version AIC and BIC will be shown in the output]
```{r}
summary(model11<-glm(relschol~white+factor(attend)+age+agesq,data=dat,family=binomial(link="logit")))
```
A. attend as continuous variable 

## 12
Run a probit regression predicting the probability attending religious school using white, attend (as continuous variable), age and agesq.
In one sentence, explain the meaning of the coefficient for "attend" in this probit model.
```{r}
summary(model12<-glm(relschol~white+attend+age+agesq,data=dat,family=binomial(link="probit")))
```
The latent variable of attending religious school is on average 0.25 standard deviation higher for those attending religious services 5 days per month than for those who do not attend

# Part II  Multinomial Regression
```{r}
setwd("~/Documents/nyu/1stGradSpring/Generalized Linear Model/dataset")
drug<-read.csv("DRUGTEST.csv")
head(drug)
```
## 1 
Reference Group: No Testing 

## 2 
BFCDAE

## 3
Based on the table above, we know that relative to no testing option, marijuana users are less likely to take the drug test (any group) than the non-users.-T

## 4 
```{r}
library(nnet)
drug$mjuser2<-relevel(factor(drug$mjuser),ref = "no")
drug$gender2<-relevel(factor(drug$gender),ref = "female")
drug$income2<-relevel(factor(drug$income),ref="low")
drug$south2<-relevel(factor(drug$south),ref = "no")
drug$construc2<-relevel(factor(drug$construc),ref="no")
drug$drugtest2<-relevel(factor(drug$drugtest),ref = "no testing program")
multi.fit<-multinom(drugtest2~mjuser2+age+educate+gender2+income2+south2+construc2,data=drug)
summary(multi.fit)
```

## 5
Based on the model from previous question, predict the probabilities of having each of the four drug testing outcomes for a male who is 30 years old, has 12 year of education, low income, comes from south, and works in the construction who has used marijuana in part year. 
```{r}
newdrug <- matrix(c("yes","male","low","yes","yes",30,12),nrow = 1,ncol = 7,byrow = T)
newdrug<-as.data.frame(newdrug)
names(newdrug)<-c("mjuser2","gender2","income2","south2","construc2","age","educate")
newdrug$age<-as.numeric(as.character(newdrug$age))
newdrug$educate<-as.numeric(as.character(newdrug$educate))
predict(multi.fit,newdrug[1,],"probs")
sum(predict(multi.fit,newdrug[1,],"probs"))
predict(multi.fit,newdrug[1,],"probs")
```

For this individual, 

the probability of no testing is  0.830

the probability of pre-employment testing & random testing is  0.098
the probability of random testing is  0.0155

the probability of pre-employment testing is  0.057














