---
title: "HW3"
author: "Charlotte Li"
date: "5/8/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("~/Documents/nyu/1stGradSpring/Generalized Linear Model/dataset")
#setwd("~/Desktop/Generalized Linear Model/dataset")
math<-read.csv("math.csv")
attach(math)
head(math)
```

# Q1 
Total number of schools and average students in each school
```{r}
#number of schools
length(unique(school))
##avg students in each school
nrow(math)/length(unique(school))
```
# Q2. Calculate the math proficiency rate (i.e. the percentage of math proficient students) in each school, and answer the following TRUE/FALSE question. Half of the schools have math proficiency rate lower than 41.59%.
```{r}
prof<-tapply(math$mathlev,math$school,sum)
stu.per.sch<-table(math$school)
table(prof/stu.per.sch<0.4159)/length(table(math$school))
## median
median(prof/stu.per.sch)
```
True

# Q3 Student level predictor 
Minoirty, Sex, ses, pracad

# Q4 School level predictor 
size, disclim, meanses, sector,pracad

# Q5 Recode
```{r}
library(plyr)
math$minority2 <- revalue(math$minority, c("Yes"="1", "No"="0"))
math$female <- revalue(math$sex, c("Female"="1", "Male"="0"))
math$public2 <- revalue(math$sector, c("Public"="1", "Catholic"="0"))
```

# Q6 Run a logistic regression including only student level predictors
```{r}
library(lme4)
Q6<-glm(mathlev~minority2+female+ses,math,family=binomial(logit))
summary(Q6)
```

# Q7 Suggest at least one way to improve the logistic regression you run in question 6 that will help us better understand the effects of various factors on students' math proficiency. Briefly explain why.
Modeling the cluster effects via random effect coefficients
A regression model for clustered data that include both the fixed effect and random is called mixed effect model. Multilevel models, random effect models, random coefficients models, hierarchical models.

# Q8 Run a random effect logistic segression with an additional school level random effect 
```{r}
Q8<-glmer(mathlev~minority2+female+ses+(1|school),math,family=binomial(logit))
summary(Q8)
```

# Q9 Variance of random effect
0.2834 

# Q10 Explain variance
The variance of the random effect explains the variability between schools.The random effect is 0.2834 suggesting it explains 0.2834 of the variance in log odds of students' math proficiency level. 

# Q11 ICC
```{r}
0.2834 /(0.2834 +(3.1415926)^2/3)
```

# Q12
```{r}
exp(Q6$coefficients)
exp(Q8@beta)
```

After exponentiate the coefficients from Q6 and Q8, the exp(betas) did not differ too much. If there is a big random effect, then the difference would be big between the two approaches. Yet, the difference is rather small. ICC is only 0.08, meaning only 8% of the variation is in school level random effect, and that is pretty small. 

#Q13 The random effect logistic regression model in question 8 has smaller AIC value than the logistic regression model in question 6.
True

# Q14 Expand Q8 model by including school level predictors 
```{r}
Q14<-glmer(mathlev~minority2+female+ses+size+public2+pracad+disclim+meanses+(1|school),math,family=binomial(logit))
summary(Q14)
```

Significant variables: minoirty, female, ses, size,pracad

## 15 The variance of the random effect in Q14 is smaller than in Q8
true













